Hybrid recurrent model for regime switching nonstationary time series

### 2021-11-25 iclr2022/hybridmodel_iclr.tex
add fina result on 10k/1k/100ts dataset

\textcolor{blue}{Result on 1000ts}  \\
%For our 3 metrics and for each model, we compute the average and standard deviation on all our sequences on the final year on the \numberts sequences. Results are shown in the Table ???.  
Table~\ref{tab:metricresults} summarize the results of all our benchmark models and our 2 hybrid methods for the 3 metrics described above. Values are errors averaged over our \numberts on the test set. For the statistical models, we can see a large domination of the TBATS model with an OWA equal to 0.847. It is one of the main reasons why we use this model on one of our hybrid frameworks as the parametric model. However, we can see that the full LSTM model \textit{lstm} is the best model over our benchmark candidates. This result shows that, with a large dataset, deep learning models can outperform the traditional approaches on average. An example of prediction of all the benchmark methods is displayed in the Figure~\ref{fig:forecastexample}.Our two different regime switching hybrid models achieve a higher level of accuracy compared to their references. For the \textit{rshm-ets}, we pass from ??? for the \textit{ets} model alone to ???? for the hybrid model. Its $\classifier$ threshold $\threshold$ was fixed at ???. With this threshold, the error-corrector model correct ???? sequences of the $\numberts$. On these corrected time series, compared to the \textit{ets} approach, our approach MASE falls from ??? to ???. However \textit{rshm-ets} doesn't succeed at outperforming some of the benchmark models, for example \textit{tbats} of \textit{lstm}. The \textit{rshm-tbats} version clearly outperforms all the previous models. Based on a \textit{tbats} model, which already achieves an interesting level of accuracy, \textit{rshm-tbats} reach the higher level with a MASE equal to 0.728 and and the lower OWA equal to 0.840. With a threshold fixed at 0.5, this version of our hybrid framework corrects 294 TBATS predictions. On these 294 trends, the \textit{tbats} has an OWA equal to 0.867 and with the regime switching model correction, we reach an OWA equal to 0.845. An example of prediction of our two regime-switching hybrid recurrent model is presented in the Figure~\ref{fig:hybridforecast}. Other additional examples and results can be found in Appendix.

\begin{figure}
  \centering
    \includegraphics[width=1.\linewidth]{figure/example_prediction}
  \caption{Example of benchmarck models's predictions on an emerging trend}
\label{fig:forecastexample}
\end{figure}

\begin{figure}
  \centering
    \includegraphics[width=1.\linewidth]{figure/rshm_tbats_prediction_example}
  \caption{Example of the \textit{rshm-ets} model prediction. In blue the ets reference and in red the $\predictor$ correction. $\classifier$ return ??? that is greater than its threshold ???: we use then the correction.}
  \centering
    \includegraphics[width=1.\linewidth]{figure/rshm_tbats_prediction_example}
  \caption{Example of the \textit{rshm-tbats} model prediction. In blue the Tbats reference and in red the $\predictor$ correction. $\classifier$ return 0.537 that is greater than its threshold 0.5: we use the RNN correction.}
\label{fig:hybridforecast}
\end{figure}

\begin{table*}
  \caption{Results summary on our Fashion dataset. For each metric, we compute the average and the standard deviation on all the test set.}
  \centering
  \begin{tabular}{lllllll}
    \hline
    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
    \hline
     \textit{snaive} & 0.871 & 0.255 & 25.29 & 14.90 & 1. & 0.\\
     \textit{ets} & 0.801 & 0.274 & 24.18 & 17.51 & 0.944 & 0.286\\
     \textit{thetam} & 0.849 & 0.364 & 24.58 & 17.07 & 0.980 & 0.333\\
     \textit{stlm} & 0.765 & 0.265 & 22.22 & 13.77 & 0.885 & 0.188\\
     \textit{tbats} & 0.734 & 0.254 & 21.19 & 13.79 & 0.847 & 0.171\\
     \textit{auto.arima} & 0. & 0. & . & . & . & 0.\\
     \textit{lstm} & 0.765 & 0.278 & 21.00 & 11.42 & 0.872 & 0.160\\
     \textit{rshm-ets} & 0. & 0. & . & . & . & 0.\\
     \textit{\textbf{rshm-tbats}} & 0.728 & 0.245 & 21.02 & 13.54 & 0.840 & 0.159\\
     \textit{oracle} & 0.716 & 0.239 & 20.74 & 13.44 & 0.828 & 0.156\\
    \hline
  \end{tabular}
\label{tab:metricresults}
\end{table*}

We also compute for each method a confusion matrix on the increase/flat/decreasing classification task. Our two hybrid models show interesting results. They achieve a better accuracy by detecting new increasing/decreasing trends or reducing the number of huge errors, predicting the opposite change direction. For example \textit{rshm-tbats} divide by 2 the number of classification errors of its reference. Another interesting result is that \textit{lstm} predicts only flat trends.

\begin{table*}
  \caption{\textit{ets} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 43 & 145 & 20 \\
    true-flat-trend & 43 & 544 & 34 \\
    true-inc-trend & 16 & 122 & 33 \\
    \hline
  \end{tabular}
  \caption{\textit{rshm-ets} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 0 & 0 & 0 \\
    true-flat-trend & 0 & 0 & 0 \\
    true-inc-trend & 0 & 0 & 0 \\
    \hline
  \end{tabular}
  \caption{\textit{rshm-ets-ws} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 0 & 0 & 0 \\
    true-flat-trend & 0 & 0 & 0 \\
    true-inc-trend & 0 & 0 & 0 \\
    \hline
  \end{tabular}
\label{tab:etsclass}
\end{table*}

\begin{table*}
  \caption{\textit{tbats} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 33 & 167 & 8 \\
    true-flat-trend & 21 & 578 & 22 \\
    true-inc-trend & 7 & 136 & 28 \\
    \hline
  \end{tabular}
  \caption{\textit{rshm-tbats} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 27 & 174 & 7 \\
    true-flat-trend & 11 & 597 & 13 \\
    true-inc-trend & 5 & 140 & 26 \\
    \hline
  \end{tabular}
  \caption{\textit{rshm-tbats-ws} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 26 & 171 & 11 \\
    true-flat-trend & 17 & 579 & 25 \\
    true-inc-trend & 3 & 136 & 32 \\
    \hline
  \end{tabular}
\label{tab:tbatsclass}
\end{table*}

\begin{table*}
  \caption{\textit{lstm} model confusion matrix}
  \centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 0 & 208 & 0 \\
    true-flat-trend & 0 & 621 & 0 \\
    true-inc-trend & 0 & 171 & 0 \\
    \hline
  \end{tabular}
  \caption{\textit{lstm-ws} model confusion matrix}
  %\centering
  \begin{tabular}{llll}
    \hline
    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
    \hline
	true-dec-trend & 1 & 205 & 2 \\
    true-flat-trend & 2 & 619 & 0 \\
    true-inc-trend & 0 & 168 & 3 \\
    \hline
  \end{tabular}
\label{tab:lstmclass}
\end{table*}

The Table~\ref{tab:metricresultsws},Table~\ref{tab:etsclass} and Table~\ref{tab:tbatsclass} presents how our candidates are able to understand and leverage our weak signals in terms of accuracy and classification. The full lstm approach shows a slight increase, from 0.872 to 0.867 for the OWA metric, and correctly predicts 4 increasing/decreasing trends. \textcolor{blue}{No for now, WIP} A real improvement is shown by our proposed regime switching framework. With a threshold fixed at 0.465, this version of our hybrid framework corrects 118 TBATS predictions. On these 118 trends,   the \textit{tbats} has an OWA equal to 0.893 and with the regime switching model correction, we reach an OWA equal to 0.868.

\begin{table*}
  \caption{Results summary on our Fashion dataset. We compare methods without our weak signal and than method but train with our weak signals as an additional input}
  \centering
  \begin{tabular}{lllllll}
    \hline
    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
    \hline
     \textit{lstm} & 0.765 & 0.278 & 21.00 & 11.42 & 0.872 & 0.160\\
     \textit{rshm-ets} & 0. & 0. & . & . & . & 0.\\
     \textit{\textbf{rshm-tbats}} & 0.728 & 0.245 & 21.02 & 13.54 & 0.840 & 0.159\\
    \hline
     \textit{lstm-ws} & 0.758 & 0.271 & 20.89 & 11.46 & 0.867 & 0.163\\
     \textit{rshm-ets-ws} & 0. & 0. & . & . & . & 0.\\
     \textit{\textbf{rshm-tbats-ws}} & 0.731 & 0.248 & 21.10 & 13.53 & 0.844 & 16.23\\
    \hline
  \end{tabular}
\label{tab:metricresultsws}
\end{table*}

>>>>

%\textcolor{blue}{Result on 1000ts}  \\
%For our 3 metrics and for each model, we compute the average and standard deviation on all our sequences on the final year on the \numberts sequences. Results are shown in the Table ???.  
%Table~\ref{tab:metricresults} summarize the results of all our benchmark models and our 2 hybrid methods for the 3 metrics described above. Values are errors averaged over our \numberts on the test set. For the statistical models, we can see a large domination of the TBATS model with an OWA equal to 0.847. It is one of the main reasons why we use this model on one of our hybrid frameworks as the parametric model. However, we can see that the full LSTM model \textit{lstm} is the best model over our benchmark candidates. This result shows that, with a large dataset, deep learning models can outperform the traditional approaches on average. An example of prediction of all the benchmark methods is displayed in the Figure~\ref{fig:forecastexample}.Our two different regime switching hybrid models achieve a higher level of accuracy compared to their references. For the \textit{rshm-ets}, we pass from ??? for the \textit{ets} model alone to ???? for the hybrid model. Its $\classifier$ threshold $\threshold$ was fixed at ???. With this threshold, the error-corrector model correct ???? sequences of the $\numberts$. On these corrected time series, compared to the \textit{ets} approach, our approach MASE falls from ??? to ???. However \textit{rshm-ets} doesn't succeed at outperforming some of the benchmark models, for example \textit{tbats} of \textit{lstm}. The \textit{rshm-tbats} version clearly outperforms all the previous models. Based on a \textit{tbats} model, which already achieves an interesting level of accuracy, \textit{rshm-tbats} reach the higher level with a MASE equal to 0.728 and and the lower OWA equal to 0.840. With a threshold fixed at 0.5, this version of our hybrid framework corrects 294 TBATS predictions. On these 294 trends, the \textit{tbats} has an OWA equal to 0.867 and with the regime switching model correction, we reach an OWA equal to 0.845. An example of prediction of our two regime-switching hybrid recurrent model is presented in the Figure~\ref{fig:hybridforecast}. Other additional examples and results can be found in Appendix.

%\begin{figure}
%  \centering
%    \includegraphics[width=1.\linewidth]{figure/example_prediction}
%  \caption{Example of benchmarck models's predictions on an emerging trend}
%\label{fig:forecastexample}
%\end{figure}

%\begin{figure}
%  \centering
%    \includegraphics[width=1.\linewidth]{figure/rshm_tbats_prediction_example}
%  \caption{Example of the \textit{rshm-ets} model prediction. In blue the ets reference and in red the $\predictor$ correction. $%\classifier$ return ??? that is greater than its threshold ???: we use then the correction.}
%  \centering
%    \includegraphics[width=1.\linewidth]{figure/rshm_tbats_prediction_example}
%  \caption{Example of the \textit{rshm-tbats} model prediction. In blue the Tbats reference and in red the $\predictor$ correction. $\classifier$ return 0.537 that is greater than its threshold 0.5: we use the RNN correction.}
%\label{fig:hybridforecast}
%\end{figure}

\begin{table*}
  \caption{Results summary on our 10000ts Fashion dataset. For each metric, we compute the average on all our time series. For our two apporaches using neural network, we trained 10 models with different seeds and we display the mean and the standard deviation of our 10 results.}
  \centering
  \begin{tabular}{lllllll}
    \hline
    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
    \hline
     \textit{snaive} & 0.881 & - & 25.27 & - & 1. & -\\
     \textit{thetam}  & 0.844 & - & 24.17 & - & 0.970 & -\\
     \textit{arima} & 0.826 & - & 23.63 & - & 0.949 & -\\
     \textit{stlm} & 0.770 & - & 22.05 & - & 0.882 & -\\
     \textit{tbats} & 0.745 & - & 21.13 & - & 0.850 & -\\
     \textit{lstm-withws} & 0.728 & 0.004 & 20.18 & 0.085 & 0.837 & 0.004\\
     \textit{lstm-nows} & 0.724 & 0.003 & 20.08 & 0.052  & 0.825 & 0.004\\
     \textit{hybrid-withws} & 0.713 & 0.005 & 20.53 & 0.170 & 0.825 & 0.009\\
     \textit{hybrid-nows} & 0.715 & 0.002 & 20.41 & 0.056 & 0.816 & 0.002\\
    \hline
  \end{tabular}
\label{tab:metricresults}
\end{table*}

\begin{table*}
  \caption{Results summary on our 1000ts Fashion dataset.For each metric, we compute the average on all our time series. For our two apporaches using neural network, we trained 10 models with different seeds and we display the mean and the standard deviation of our 10 results.}
  \centering
  \begin{tabular}{lllllll}
    \hline
    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
    \hline
     \textit{snaive} & 0.871 & - & 25.29 & - & 1. & -\\
     \textit{thetam}  & 0.849 & - & 24.58 & - & 0.980 & -\\
     \textit{arima} & 0.821 & - & 23.69 & - & 0.951 & -\\
     \textit{stlm} & 0.765 & - & 22.22 & - & 0.885 & -\\
     \textit{lstm-nows} & 0.740 & 0.007 & 20.56 & 0.154 & 0.856 & 0.008\\
     \textit{tbats} & 0.734 & - & 21.19 & - & 0.847 & -\\
     \textit{hybrid-nows} & 0.719 & 0.002 & 20.79 & 0.094 & 0.833 & 0.004\\
    \hline
  \end{tabular}
\label{tab:metricresults}
\end{table*}

\begin{table*}
  \caption{Results summary on our 100ts Fashion dataset. For each metric, we compute the average on all our time series. For our two apporaches using neural network, we trained 10 models with different seeds and we display the mean and the standard deviation of our 10 results.}
  \centering
  \begin{tabular}{lllllll}
    \hline
    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
    \hline
     \textit{snaive} & 0.876 & - & 26.98 & - & 1. & -\\
     \textit{thetam}  & 0.823 & - & 25.11 & - & 0.944 & -\\
     \textit{arima} & 0.814 & - & 24.61 & - & 0.940 & -\\
     \textit{lstm-nows} & 0.767 & 0.045 & 22.22 & 0.848 & 0.877 & 0.051\\
     \textit{stlm} & 0.742 & - & 23.03 & - & 0.863 & -\\
     \textit{tbats} & 0.745 & - & 22.16 & - & 0.856 & -\\
     \textit{hybrid-nows} & 0.739 & 0.003 & 22.31 & 0.083 & 0.845 & 0.003\\
    \hline
  \end{tabular}
\label{tab:metricresults}
\end{table*}

%We also compute for each method a confusion matrix on the increase/flat/decreasing classification task. Our two hybrid models show interesting results. They achieve a better accuracy by detecting new increasing/decreasing trends or reducing the number of huge errors, predicting the opposite change direction. For example \textit{rshm-tbats} divide by 2 the number of classification errors of its reference. Another interesting result is that \textit{lstm} predicts only flat trends.

%\begin{table*}
%  \caption{\textit{ets} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 43 & 145 & 20 \\
%    true-flat-trend & 43 & 544 & 34 \\
%    true-inc-trend & 16 & 122 & 33 \\
%    \hline
%  \end{tabular}
%  \caption{\textit{rshm-ets} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 0 & 0 & 0 \\
%    true-flat-trend & 0 & 0 & 0 \\
%    true-inc-trend & 0 & 0 & 0 \\
%    \hline
%  \end{tabular}
%  \caption{\textit{rshm-ets-ws} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 0 & 0 & 0 \\
%    true-flat-trend & 0 & 0 & 0 \\
%    true-inc-trend & 0 & 0 & 0 \\
%    \hline
%  \end{tabular}
%\label{tab:etsclass}
%\end{table*}

%\begin{table*}
%  \caption{\textit{tbats} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 33 & 167 & 8 \\
%    true-flat-trend & 21 & 578 & 22 \\
%    true-inc-trend & 7 & 136 & 28 \\
%    \hline
%  \end{tabular}
%  \caption{\textit{rshm-tbats} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 27 & 174 & 7 \\
%    true-flat-trend & 11 & 597 & 13 \\
%    true-inc-trend & 5 & 140 & 26 \\
%    \hline
%  \end{tabular}
%  \caption{\textit{rshm-tbats-ws} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 26 & 171 & 11 \\
%    true-flat-trend & 17 & 579 & 25 \\
%    true-inc-trend & 3 & 136 & 32 \\
%    \hline
%  \end{tabular}
%\label{tab:tbatsclass}
%\end{table*}

%\begin{table*}
%  \caption{\textit{lstm} model confusion matrix}
%  \centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 0 & 208 & 0 \\
%    true-flat-trend & 0 & 621 & 0 \\
%    true-inc-trend & 0 & 171 & 0 \\
%    \hline
%  \end{tabular}
%  \caption{\textit{lstm-ws} model confusion matrix}
%  %\centering
%  \begin{tabular}{llll}
%    \hline
%    &  dec-trend-pred  & flat-trend-pred & inc-trend-pred  \\
%    \hline
%	true-dec-trend & 1 & 205 & 2 \\
%    true-flat-trend & 2 & 619 & 0 \\
%    true-inc-trend & 0 & 168 & 3 \\
%    \hline
%  \end{tabular}
%\label{tab:lstmclass}
%\end{table*}

%The Table~\ref{tab:metricresultsws},Table~\ref{tab:etsclass} and Table~\ref{tab:tbatsclass} presents how our candidates are able to understand and leverage our weak signals in terms of accuracy and classification. The full lstm approach shows a slight increase, from 0.872 to 0.867 for the OWA metric, and correctly predicts 4 increasing/decreasing trends. \textcolor{blue}{No for now, WIP} A real improvement is shown by our proposed regime switching framework. With a threshold fixed at 0.465, this version of our hybrid framework corrects 118 TBATS predictions. On these 118 trends,   the \textit{tbats} has an OWA equal to 0.893 and with the regime switching model correction, we reach an OWA equal to 0.868.

%\begin{table*}
%  \caption{Results summary on our Fashion dataset. We compare methods without our weak signal and than method but train with our weak signals as an additional input}
%  \centering
%  \begin{tabular}{lllllll}
%    \hline
%    &  MASE-mean  & MASE-std & SMAPE-mean & SMAPE-std & OWA-mean & OWA-std  \\
%    \hline
%     \textit{lstm} & 0.765 & 0.278 & 21.00 & 11.42 & 0.872 & 0.160\\
%     \textit{rshm-ets} & 0. & 0. & . & . & . & 0.\\
%     \textit{\textbf{rshm-tbats}} & 0.728 & 0.245 & 21.02 & 13.54 & 0.840 & 0.159\\
%    \hline
%     \textit{lstm-ws} & 0.758 & 0.271 & 20.89 & 11.46 & 0.867 & 0.163\\
%     \textit{rshm-ets-ws} & 0. & 0. & . & . & . & 0.\\
%     \textit{\textbf{rshm-tbats-ws}} & 0.731 & 0.248 & 21.10 & 13.53 & 0.844 & 16.23\\
%    \hline
%  \end{tabular}
%\label{tab:metricresultsws}
%\end{table*}
